---
title: "Frontiers manuscript figures and tables"
author: "dan.auerbach@dfw.wa.gov"
date: "`r Sys.Date()`"
format:
  docx:
    reference-doc: /Users/auerbdaa/O/code/quarto_template.docx
  html:
    embed-resources: true
    theme: yeti 
    code-fold: true
    toc: true
    toc-location: left
    grid:
      sidebar-width: 180px
      body-width: 1100px
      margin-width: 20px
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 9, fig.height = 10)

library("tidyverse", quietly = T)
library("sf")
library("patchwork")
library("gt")
theme_set(theme_minimal()) 

# dir_data_common <- "~/T/DFW-Team WDFW Watershed Synthesis - data_common"
# epsg <- 2927 #WA state standard; NAD83(HARN)/ft

load("~/T/DFW-Team WDFW Watershed Synthesis - IMW analyses/imw_hab_250307.RData")

```

# Abstract

# Introduction

Widespread recognition of declining Pacific salmon populations has led to major, often publicly-funded, recovery efforts (Bilby et al. 2024). Protection and rehabilitation of the freshwater and estuarine systems where salmon spawn and rear is essential to these efforts, and, in 1999, the Washington State legislature created the Salmon Recovery Funding Board (SRFB) and a system of regional recovery organizations to distribute resources for projects meant to benefit fish habitat. The SRFB subsequently requested a study to assess the efficacy of stream habitat restoration and enhancement at the "watershed scale", based on the idea that sub-populations of each salmon species complete their freshwater life histories at the spatial extent of river networks rather than that of individual projects. Despite recognition that fish integrate complex hydrologic and geomorphic processes throughout riverscapes and the fact that many monitoring and management activities occur at this scale (Roni 2005), the logistical difficulty of experimental manipulation over entire catchments and the urgency of intervention has greatly limited the evaluation of restoration practice in this way.

The Washington Intensively Monitored Watersheds (IMW) study is a collaborative, long-term, large-scale effort to meet the need for such evidence, with a study design that pairs measures of habitat conditions with salmon survival and productivity monitoring throughout multiple entire basins in several geographic "complexes" of replicate treatment and adjacent control streams (Bilby et al. 2005, Bennett et al. 2016). The inclusion of untreated sites and entire reference streams subject to comparable climate and management regimes enables strong inference regarding attribution and efficacy (Parker and Wiens 2005), and the results of this study were envisioned to directly inform the organizational processes and procedures through which various entities identify, design, propose, and implement recovery projects.

This paper describes patterns of variation in the stream habitat measures collected annually since 2007 at `r nrow(site_meta)` sites distributed within 10 streams comprising 3 western Washington complexes. These observations constitue a substantial body of empirical evidence regarding the long-term dynamics of channel form in forested catchments of northwestern North America, and they contribute a unique and valuable perspective to discussions regarding the strengths and limitations of salmon recovery efforts to date (Bilby et al. 2024). Most robustly, this large dataset reinforces the critical need to plan for spatiotemporal variation in process-based river restoration. These findings also indicate the practical challenges of detecting returns on restoration investments at the whole-stream scale, and the IMW habitat monitoring data suggest how the sociopolitical processes that currently shape restoration resource allocation may unintentionally increase this detection problem. 


# Methods

## Study area

Stream habitat data in the western Washington IMW study are collected at sites within three complexes of 10 small, adjacent watersheds (Figure 1, Table 1). The Strait of Juan de Fuca complex (ST) includes East Twin, West Twin (reference), and Deep creeks. The Hood Canal complex (HC) includes Little Anderson, Big Beef, Seabeck and Stavis (reference) creeks. The Lower Columbia complex (LC) includes Germany, Abernathy, and Mill (reference) creeks. These smaller watersheds were selected with the intent of increasing the proportion of stream networks that experienced restoration "treatment" within the study duration and to increase the sampling density for monitoring surveys.  

```{r figMAP_stream_sites_placeholder, fig.cap="Western Washington IMW stream habitat monitoring sites for Hood Canal (a,d), Lower Columbia (b,e), and Strait of Juan de Fuca (c,f) complexes. Stavis, Mill and West Twin creeks were designated as control watersheds in original study design."}

bsmp_str <- map(
  set_names(unique(sf_site_meta$cmplx_desc))
  ,
  ~maptiles::get_tiles(
    st_as_sfc(st_bbox(sf_site_meta |> filter(cmplx_desc == .x))) |> st_buffer(5000),
    #provider = "OpenStreetMap",
    provider = "Esri.WorldStreetMap",
    crop = TRUE
  )
)

bsmp_img <- map(
  set_names(unique(sf_site_meta$cmplx_desc))
  ,
  ~maptiles::get_tiles(
    st_as_sfc(st_bbox(sf_site_meta |> filter(cmplx_desc == .x))) |> st_buffer(5000),
    provider = "Esri.WorldImagery",
    crop = TRUE
  )
)

bsmp_top <- map(
  set_names(unique(sf_site_meta$cmplx_desc))
  ,
  ~maptiles::get_tiles(
    st_as_sfc(st_bbox(sf_site_meta |> filter(cmplx_desc == .x))) |> st_buffer(5000),
    provider = "OpenTopoMap",
    crop = TRUE
  )
)

wrap_plots(map(
  unique(sf_site_meta$cmplx_desc)
  ,
  ~ggplot() + 
    tidyterra::geom_spatraster_rgb(data = bsmp_str[[.x]]) +
    geom_sf(data = sf_site_meta |> filter(cmplx_desc == .x), aes(color = strm), show.legend = F) +
    geom_sf_text(
      data = sf_site_meta |> 
        filter(cmplx_desc == .x) |> 
        #mutate(strm = as.character(strm)) |> 
        group_by(strm) |> summarize() |> st_centroid(),
      mapping = aes(label = strm, color = strm), show.legend = F) +
    scale_color_manual(values = pal_strm) +
    ggspatial::annotation_scale() +
    theme_void()
), ncol = 1) -
  wrap_plots(map(
  unique(sf_site_meta$cmplx_desc)
  ,
  ~ggplot() + 
    tidyterra::geom_spatraster_rgb(data = bsmp_img[[.x]]) +
    geom_sf(data = sf_site_meta |> filter(cmplx_desc == .x), color = "orange") +
    ggspatial::annotation_scale() +
    theme_void()
), ncol = 1) +
  plot_annotation(tag_levels = "a")

#   wrap_plots(map(
#   unique(sf_site_meta$cmplx_desc)
#   ,
#   ~ggplot() + 
#     tidyterra::geom_spatraster_rgb(data = bsmp_top[[.x]]) +
#     geom_sf(data = sf_site_meta |> filter(cmplx_desc == .x), color = "orange") +
#     ggspatial::annotation_scale() +
#     theme_void()
# ), ncol = 1) 

```

```{r tabStrmMetadata}
site_meta |> 
  summarise(
    n_site = n(),
    across(year_n, median),
    across(n_tsct_tot, sum),
    .by = c(cmplx_desc, strm)
  ) |>
  mutate(
    `Treatment/Control` = if_else(str_detect(strm, "Stavis|Mill|West"), "control", "treatment")
    ,colfill = alpha(pal_strm[strm], 0.7)
    ) |> 
  select(cmplx_desc, strm, colfill, `Treatment/Control`,
         Sites = n_site, `Median years/site` = year_n, `Total transects` = n_tsct_tot) |> 
  gt(groupname_col = "cmplx_desc", rowname_col = "strm") |> 
  cols_hide(columns = "colfill") |> 
  fmt_number(columns = where(is.numeric), decimals = 0) |> 
  tab_style(
    style = cell_fill(color = from_column("colfill")), locations = cells_body()
  )

```

Intensive annual habitat surveys were based on Lazorchak et al. (1998), Crawford et al. (2002), and Roper et al. (2003). Survey protocols are similar to those used by the Washington Watershed Health Monitoring Program and the Oregon Coastal Coho Surveys (Anlauf et al. 2009). Habitat sampling locations (hereafter sites) were identified using a random, spatially balanced design stratified by stream order (Stevens and Olsen 2004). Surveys were conducted between June and October, beginning in 2004 and ongoing, with rotation among watersheds approximately weekly to minimize seasonal bias. Given the interest in detecting interannual trends, and following preliminary analyses of survey data collected in 2004 – 2006, the approach was modified to resurvey a consistent set of sites in each watershed. Only data from surveys conducted in 2007 and onward are presented. Some sites have been added following the random, spatially balanced design due to changes in accessibility.

At each site, measurements and counts are made at and between 21 equally spaced cross-sectional transects, positioned equidistant and perpendicular to flow along a length of stream that is the longer of either 40 bankfull widths or 300m. Measurements at each transect included bankfull width (to the nearest 0.0 m), water depth (≥5 points to the nearest 0.00 m), and tallies of substrate size classes. Wood within the bankfull channel was counted by length and diameter classes between transects. Substrate was classified at approximately 5 points per transect as either smooth bedrock or rough bedrock (> 4 m), boulder (250 – 4000 mm), cobble (64 – 250 mm), coarse gravel (16 – 64 mm), fine gravel (2 – 16 mm), sand (0.06 – 2 mm), fines (< 2 mm), hardpan, wood, or other. When found, dry channels were sampled and all possible measurements were made (e.g., depth is not measured).

## Data and analyses

```{r emap_pull, echo=FALSE, eval=FALSE}
#declare 'con_emap' odbc::dbConnect
library(tidyverse)

emap <- list()

# rows per year, site metadata: location, length, etc.
# limiting to sampled, 2007 forward
# replacing -9999s with NA
emap$reaches_sampled <- tbl(
  con_emap,
  dbplyr::in_schema("dbo", "ZCORE_Reaches_All_V")
  ) |>
  filter(
    Sampl_Or_Not_Ind == 1,
    Srvy_Yr > 2006
    ) |>
  select(
    cmplx_abrv = COMPLEX_NAME_Abrvd,
    cmplx_desc = COMPLEX_NAME_Desc,
    strm = STREAM_NAME,
    site = Site_Name,
    year = Srvy_Yr,
    Reach_Site_Id,
    Reach_Survey_Id,
    x_lat = Xsite_Lat_Coord,
    x_lon = Xsite_Lngtd_Coord,
    smpl_mthd_desc = SAMPL_METHOD_Desc,
    rch_lgth_meas_desc = REACH_LNGTH_MEAS_LOC_Desc,
    Xsite_Distnc_Up_Meas, Xsite_Distnc_Down_Meas
    #STREAM_NAME_LLID_100_Code, Total_Reach_Lgth, Move_Reach_Ind
  ) |>
  collect() |>
  rename_with(tolower) |>
  mutate(
    across(
      c(xsite_distnc_up_meas, xsite_distnc_down_meas),
      ~if_else(. < 0, NA_integer_, .)
    ),
    up_down_lgth = xsite_distnc_up_meas + xsite_distnc_down_meas
  )


#bankfull, wetted width, thalweg depth, size channels, bars, etc.
#usually ~21 transects per site per year
#limited to 2007 forward, replacing -9999 with NAs
#dropping 'code' LU fields, 'verify' date & time, crewname
emap$transects <- tbl(
  con_emap,
  dbplyr::in_schema("dbo", "ZCORE_Transects_V")
  ) |>
  filter(
    Srvy_Yr > 2006
  ) |>
  select(
    cmplx_desc = COMPLEX_NAME_Desc,
    strm = STREAM_NAME,
    site = Site_Name,
    year = Srvy_Yr,
    Reach_Survey_Id,
    x_lat = Xsite_Lat_Coord,
    x_lon = Xsite_Lngtd_Coord,
    Transect_Id, Transect_Letter,
    Transect_Extra_Side_Channel_Ind,
    Transect_Fish_Present_Ind,
    Transect_Wet_Wdth_Meas,
    Transect_Bar_Wdth_Present_Ind, Transect_Bar_Wdth_Meas,
    Transect_Bankfull_Wdth_Meas, Transect_Bankfull_Hgt_Meas,
    Transect_Thalweg_Dpth_Meas,
    CHANNEL_UNIT_CODE_Abrvd, CHANNEL_UNIT_CODE_Desc, #,Transect_Channel_Unit_Code,
    POOL_FORM_CODE_Desc, #,Transect_Pool_Form_Code,
    Transect_Backwater_Present_Ind,
    Pool_Non_Pool_Desc #,Pool_Non_Pool_Code,
  ) |>
  collect() |>
  rename_with(tolower) |>
  drop_na(transect_id) |>
  mutate(
    across(
      c(ends_with("_meas")),
      ~if_else(. < 0, NA_real_, .)
    )
  )

#usually 5 substrate measures per transect, ~21 transects per site per year
#limited to 2007 forward, replacing -9999 with NAs
emap$substrate <- tbl(
  con_emap,
  dbplyr::in_schema("dbo", "ZCORE_Transects_Substrate_V")
  ) |>
  filter(Srvy_Yr > 2006) |>
  select(
    cmplx_desc = COMPLEX_NAME_Desc,
    strm = STREAM_NAME,
    site = Site_Name,
    year = Srvy_Yr,
    Reach_Survey_Id,
    Transect_Id, Transect_Letter,
    CHANNEL_UNIT_CODE_Abrvd, #,Transect_Channel_Unit_Code,
    Transect_Substrate_Id,
    TRANSECT_SUBSTRATE_SAMPL_PT_Abrvd, #,TRANSECT_SUBSTRATE_SAMPL_PT_Code,
    Transect_Substrate_Sampl_Dpth_Meas,
    TRANSECT_SUBSTRATE_SIZE_CLAS_Abrvd, TRANSECT_SUBSTRATE_SIZE_CLAS_Desc #,TRANSECT_SUBSTRATE_SIZE_CLAS_Code
  ) |>
  collect() |>
  rename_with(tolower) |>
  mutate(
    across(
      c(ends_with("_meas")),
      ~if_else(. < 0, NA_real_, .)
    )
  )

#measured for 'segments' between (x-section) transects
#counts for multiple length and diameter classes, in-channel and 'bridging'
#no coercion to NA (yet)
emap$lwd <- tbl(
  con_emap,
  dbplyr::in_schema("dbo", "ZCORE_ReachSegments_LWD_V")
  ) |> #colnames()
  filter(Srvy_Yr > 2006) |>
  select(
    cmplx_desc = COMPLEX_NAME_Desc,
    strm = STREAM_NAME,
    site = Site_Name,
    year = Srvy_Yr,
    Reach_Survey_Id,
    everything()
  ) |>
  collect() |>
  drop_na(LWD_CLAS_Code) |>
  rename_with(~str_remove(.,"LWD_CLAS_"), contains("LWD_CLAS_")) |>
  rename_with(tolower)

glimpse(emap)

DBI::dbDisconnect(con_emap)
rm(uid, pwd, con_emap)
saveRDS(emap, "emap_pull.rds")

```

```{r build_envi, echo=FALSE, eval=FALSE}
#this chunk rebuilds the environment from the full query object
#  'emap' list of only raw/orig query (before raw and wrangled)
#  'hab' list of wrangled objects
#!!hab object here is NOT same as in high_var_stream_hab chunk 'build_hab'!!
# that chunk high_var_stream_hab>>build_hab has more comments on qaqc and wrangling checks

#file.info("~/O/code/imw/emap_pull.rds")['mtime']
file.info("~/T/DFW-Team WDFW Watershed Synthesis - IMW analyses/emap_pull.rds")['mtime']
emap <- readRDS("~/T/DFW-Team WDFW Watershed Synthesis - IMW analyses/emap_pull.rds")
names(emap)

#plotting palettes: W to E within complex
pal_strm <- set_names(
  c(c("#3A270A", "#D3BEAF", "#965127", "#D2A554"), #"#E4C22B" 
    c("#1E212F", "#48628A", "#A4BADF"), 
    c("#254029", "#516F25", "#94AA3D")
    ),
  c('Stavis Creek*','Seabeck Creek','Big Beef Creek','Little Anderson Creek',
    'Mill Creek*','Abernathy Creek','Germany Creek',
    'Deep Creek','West Twin River*','East Twin River'))
pal_cmplx_strm <- set_names(
  pal_strm,
  c('HC_Stavis Creek*','HC_Seabeck Creek','HC_Big Beef Creek','HC_Little Anderson Creek',
    'LC_Mill Creek*','LC_Abernathy Creek','LC_Germany Creek',
    'ST_Deep Creek','ST_West Twin River*','ST_East Twin River'))

# #leaves culverts, falls, fishladders etc.
# emap$transects |> count(channel_unit_code_abrvd, channel_unit_code_desc) |> print(n=100)
transects_excluded <- c(
  "-9999", #NAs
  "IA", #~350 inaccessible
  "OT" #10 other
  ) 


hab <- list()

#### 1. location and monitoring duration metadata
hab$site_meta <- emap$reaches_sampled |>
  summarise(
    across(c(year), list(min = min, max = max)),
    year_diff = year_max - year_min + 1,
    year_n = n(),
    across(c(x_lon, x_lat), list(mean = ~mean(., na.rm = T))),
    .by = c("cmplx_abrv", "cmplx_desc", "strm", "site")
  ) |>
  mutate(
    cmplx_desc = str_replace(cmplx_desc, "Straits", "Strait"),
    strm = if_else(
      str_detect(strm, "Stavis|Mill|West"),
      paste0(strm, "*"), 
      strm) |> 
      factor(levels = names(pal_strm))
  ) |> 
  unite("cmplx_strm", cmplx_abrv, strm, remove=F) |> 
  mutate(
    cmplx_strm = factor(cmplx_strm, levels = names(pal_cmplx_strm)),
    resto_add = as.integer(str_sub(site,4,6)) > 899
  ) |> 
  arrange(cmplx_desc, strm, year_min, site) 

#### 2. clean transects, preserving 'true' supplementals and reducing to only obs used 
#primary cross-sectional transects are A:K, with intervening A2, B2
#supplementary X-Y-Z transects at the same stream-longitudinal position can occur when field crew encounter 'substantial' flow splits
#the distinct() removes ~50 cases in HC from 2007 and 2008 of transect rows that are "X|Y|Z" lettered but are exact duplicates bankfull, wetted and depth measures
hab$transects <- emap$transects |> 
  filter(!(channel_unit_code_abrvd %in% transects_excluded)) |> 
  mutate(
    cmplx_desc = str_replace(cmplx_desc, "Straits", "Strait"),
    strm = if_else(
      str_detect(strm, "Stavis|Mill|West"),
      paste0(strm, "*"), 
      strm) |> 
      factor(levels = names(pal_strm)),
    tsct_lett = str_remove_all(transect_letter, "X|Y|Z") #but retaining orig 'transect_letter'
    ) |>
  distinct(
    cmplx_desc, strm, site, year, reach_survey_id, tsct_lett,
    pool_non_pool_desc, channel_unit_code_abrvd, channel_unit_code_desc,
    transect_bankfull_wdth_meas, transect_wet_wdth_meas, transect_thalweg_dpth_meas,
    .keep_all = T
    ) |> 
  #dropping bar indicator and width
  left_join(hab$site_meta |> distinct(strm, cmplx_strm), by = "strm") |> 
  select(
    cmplx_strm, cmplx_desc, strm, site, year, reach_survey_id, x_lon, x_lat, transect_id,
    tsct_lett, transect_letter,
    side_chan = transect_extra_side_channel_ind,
    fish_prst = transect_fish_present_ind,
    bnk_wdth = transect_bankfull_wdth_meas,
    bnk_hght = transect_bankfull_hgt_meas,
    wet_wdth = transect_wet_wdth_meas,
    thl_dpth = transect_thalweg_dpth_meas,
    pool_non_pool_desc, 
    channel_unit_code_abrvd, channel_unit_code_desc
    )

#### 3. summarize site-year channel form measures and substrate
# - number/pct of 'supplemental' transects for parallel/anastomosed 'side channels' at same longitudinal cross-section
# - number/pct of intersected (categorical) pools
# - mean of (thalweg) depth of transects intersecting pools
# - pct_gravels, note this returns fewer rows than the objects based directly on hab$transects, some strm-site-year are NA esp with filter to GC/GF
# - summarize site-year sample distrib of (main+supp) bankfull width 
#   - first need to add main and supplemental widths
#   - here not distinguishing by categorical pool_non_pool or channel_unit_code to give 'full width' regardless of different 'form types'
#   - reduces to max of 21 obs per site-year at 'base letter' positions (which can be fewer than the site-year count of all transects)
# - summarize W:D across transects after first calc'ing for supplementary (XA, YA, etc.) as distinct from primary
hab$transects_site_year <- hab$transects |> 
  summarise(
    n_tsct = n(), #same as: length(transect_letter)
    n_supp = sum(str_detect(transect_letter, "X|Y|Z")),
    pct_supp = n_supp / n_tsct,
    n_pools = sum(pool_non_pool_desc == "Pool"), #includees 'Beaver Pond'
    pct_pools = n_pools / n_tsct,
    .by = c("cmplx_strm", "strm", "site", "year", "reach_survey_id")
  ) |> 
  full_join(
    hab$transects |> 
      filter(pool_non_pool_desc == "Pool") |> 
      summarise(
        pool_depth_mean = mean(thl_dpth, na.rm = T),
        .by = c("strm", "site", "year", "reach_survey_id")
      ) 
    ,
    by = c("strm", "site", "year", "reach_survey_id")
  ) |> 
  full_join(
    emap$substrate |> 
      mutate(
        strm = if_else(
          str_detect(strm, "Stavis|Mill|West"),
          paste0(strm, "*"), 
          strm) |> 
          factor(levels = names(pal_strm))
      ) |> 
      count(strm, site, year, reach_survey_id,
            transect_substrate_size_clas_abrvd, 
            transect_substrate_size_clas_desc) |> 
      group_by(strm, site, year, reach_survey_id) |>
      mutate(n_pct = n / sum(n)) |> 
      filter(transect_substrate_size_clas_abrvd %in% c("GC", "GF")) |> 
      summarise(pct_gravel = sum(n_pct), .groups = "drop"), 
    by = c("strm", "site", "year", "reach_survey_id")
  ) |> 
  #now add width; bringing back 'wetted'
  full_join(
    hab$transects |> 
      #first sum per cross-sectional transect position, doing nothing for many non-split instances
      summarise(
        across(c(bnk_wdth, wet_wdth), sum), 
        .by = c("strm", "site", "year", "reach_survey_id", "tsct_lett")
      ) |> 
      #then aggregate to site-year distributional measures across transects
      summarise(
        across(
          c(bnk_wdth, wet_wdth),
          list(
            mean = ~mean(., na.rm = T),
            sd = ~sd(., na.rm = T),
            min = ~min(., na.rm = T),
            #med = ~median(., na.rm = T),
            max = ~max(., na.rm = T)
            #can easily calc later, leaving off until known pref to calc here
            #,trsct_cv = ~sd(., na.rm =T) / mean(., na.rm = T)
          )),
        .by = c("strm", "site", "year", "reach_survey_id")
      ), 
    by = c("strm", "site", "year", "reach_survey_id")
  ) |>  
  #now add width:depth
  #shift to wetted-thalweg, bank-bank following call with KK/WD/KS
  #summarization throws warnings whenever grouped vector is empty
  full_join(
    hab$transects |> 
      #calc W:D at every transect, including lateral 'supplementary'
      #plenty of Inf and NA
      mutate(
        wet_w_thl_d = wet_wdth / thl_dpth,
        bnk_w_bnk_d = bnk_wdth / bnk_hght
      ) |> 
      #then aggregate to site-year distributional measures across transects
      summarise(
        across(
          c(wet_w_thl_d, bnk_w_bnk_d),
          list(
            mean = ~mean(., na.rm = T),
            sd = ~sd(., na.rm = T),
            min = ~min(., na.rm = T),
            max = ~max(., na.rm = T)
          )),
        .by = c("strm", "site", "year", "reach_survey_id")
      ) 
    ,
    by = c("strm", "site", "year", "reach_survey_id")
  ) 

#### 4. summarize site-year large wood densities
#per transect, 3 length and 4 diameter classes for 2 position bridging & inchan
#always recording counts within each crossing level, including 0s
#past practice to drop smallest and shortest then sum counts over other levels
# emap$lwd |> count(diameter_code, diameter_desc, lgth_code, lgth_desc)
#    diameter_code diameter_desc lgth_code lgth_desc     n
#            <int> <chr>             <int> <chr>     <int>
#  1            10 0.1 - 0.3 m          10 1.5 - 5 m 86180
#  2            10 0.1 - 0.3 m          20 5 - 15 m  86180
#  3            10 0.1 - 0.3 m          30 >15 m     86180
#  4            20 0.3 - 0.6 m          10 1.5 - 5 m 86180
#  5            20 0.3 - 0.6 m          20 5 - 15 m  86180
#  6            20 0.3 - 0.6 m          30 >15 m     86180
#  7            30 0.6 - 0.8 m          10 1.5 - 5 m 86180
#  8            30 0.6 - 0.8 m          20 5 - 15 m  86180
#  9            30 0.6 - 0.8 m          30 >15 m     86180
# 10            40 >0.8 m               10 1.5 - 5 m 86180
# 11            40 >0.8 m               20 5 - 15 m  86180
# 12            40 >0.8 m               30 >15 m     86180
#here:
# - combining 'AlI/Part In' and 'Bridged Above'
# - aggregate over segments per stream-year
# - join site-year lengths calc'd from the up+down lengths
#    - NOW no missing following WD/KS additional qaqc: emap$reaches_sampled |> filter(is.na(up_down_lgth))
# - and NOW join site-year mean bankfull width to get areal density
#    - introducing error with "area = total_length * average_width", presumably greater with more sinuosity?
#       but preferably simpler than attempting trapezoidal calc from top/bottom transect & per-segement lengths
#       not even clear that per-segment length from 'compass-distance' table would be suitable?
# - retaining 'all' as well as larger classes; pre-2024 filtered diameter_code & lgth_code > 10, #more than 0.3m dia and 5m long
hab$lwd_site_year <- emap$lwd |>
  #count(diameter_code, diameter_desc, lgth_code, lgth_desc)
  #coerce -9999 to NA
  mutate(
    strm = if_else(
      str_detect(strm, "Stavis|Mill|West"),
      paste0(strm, "*"), 
      strm) |> 
      factor(levels = names(pal_strm)),
    lwd_cnt = if_else(lwd_cnt < 0, NA_integer_, lwd_cnt)
    ,
    diameter_code = paste0("d", diameter_code),
    lgth_code = paste0("l", lgth_code)
    ) |> 
  unite("diam_lgth", diameter_code, lgth_code) |> 
  summarise(
    #n_seg = length(reach_segment_id),
    tot_lwd_cnt = sum(lwd_cnt, na.rm=T)
    , .by = c("strm", "site", "year", "reach_survey_id",
              "diam_lgth" #"diameter_code", "lgth_code"
              )
  ) |>
  pivot_wider(names_from = diam_lgth, values_from = tot_lwd_cnt) |> 
  left_join(
    emap$reaches_sampled |> 
      select(site, year, reach_survey_id, up_down_lgth)
    , by = c("site", "year", "reach_survey_id")
  ) |> 
  left_join(
    hab$transects_site_year |> 
      select(site, year, reach_survey_id, bnk_wdth_mean)
    , by = c("site", "year", "reach_survey_id")
  ) |> 
  mutate(
    lwd_all = d10_l10 + d10_l20 + d10_l30 + 
              d20_l10 + d20_l20 + d20_l30 + 
              d30_l10 + d30_l20 + d30_l30 + 
              d40_l10 + d40_l20 + d40_l30
    ,
    lwd_no10 = d20_l20 + d20_l30 + 
               d30_l20 + d30_l30 + 
               d40_l20 + d40_l30
    ,
    lgth_wdth = up_down_lgth * bnk_wdth_mean
    ,
    across(starts_with("lwd"),
           list(
             per100 = ~100 * (. / up_down_lgth)
             ,
             per_m2 = ~(. / lgth_wdth)
             ))
    ,
    #using lower bounds of size classes for more conservative estimate?
    vlwd_vol = d20_l20 * (pi * 0.15^2 * 5) + 
      d20_l30 * (pi * 0.15^2 * 15) + 
      d30_l20 * (pi * 0.3^2 * 5) + 
      d30_l30 * (pi * 0.3^2 * 15) + 
      d40_l20 * (pi * 0.4^2 * 5) + 
      d40_l30 * (pi * 0.4^2 * 15)
  ) |>
  #exclusions per WD/KS: "QC_RchMeasXsiteDstVsCmpDst_Extrc_Null_XsiteDistances.xlsx"
  filter(
    !(year == 2019 & site %in% c(
      "DEW017","DEW034","DEW073","DEW097","DEW113","DEW125",
      "DEW273","DEW287","DEW317","DEW477","DEW026","DEW044",
      "DEW067","DEW110","DEW158","DEW170","DEW630","DEW014",
      "DEW027","DEW057","DEW080","DEW322"
    ))
  )


#### 5. sites to include
hab$incl_sites <- hab$site_meta |> #287
  #filter(year_n >= 10) #251
  filter(year_n > 3 & year_n/year_diff > 0.5) |> arrange(year_n) |> #print(n=20) #269, adds mostly newer/recent but consistent sites
  select(site)
# #excluded
# emap$reaches_sampled |>
#   summarise(
#     across(c(year), list(min = min, max = max)),
#     year_diff = year_max - year_min + 1,
#     year_n = n(),
#     .by = c("cmplx_abrv", "cmplx_desc", "strm", "site")
#   ) |> 
#   anti_join(
#     hab$incl_sites, by = "site"
#   ) |> 
#   arrange(cmplx_desc, year_n)

hab$site_meta |> count(year_n) |> arrange(desc(year_n))

#### 6. summarize measures across years per site and add back to site_meta
#first filtering to included sites
#then expanding to complete years per site-stream for explicit missing vals
#coercing Inf to NA (NaN already gets treated as NA)
#warnings are legit due to missing vals and/or irrelevant due dropping sites with few obs in primary semi_join
#preserves only mean-over-years of within-year mean-over-transects
#could add back median-over-years of within-year median-over-transects but examined diffs were qualitatively insignif
hab$site_meta <- hab$site_meta |> 
  semi_join(hab$incl_sites, by = "site") |> 
  left_join(
    hab$transects_site_year |> 
      #semi_join not needed due to left_join but could uncomment for intermediate inspection: semi_join(hab$incl_sites, by = "site") |> 
      group_by(site) |> 
      complete(year = min(hab$site_meta$year_min):max(hab$site_meta$year_max)) |> 
      ungroup() |> #glimpse()
      mutate(across(where(is.numeric), \(x){ x[is.infinite(x)] <- NA; return(x)})) |> 
      summarise(
        across(c(n_tsct, n_supp), list(tot = ~sum(.,na.rm = T))),
        pct_supp = n_supp_tot / n_tsct_tot,
        across(
          c(
            bankfull_width = bnk_wdth_mean,
            wetted_width = wet_wdth_mean,
            bankfull_W_D = bnk_w_bnk_d_mean,
            wetted_W_D = wet_w_thl_d_mean,
            pct_gravel, 
            pct_pools, 
            pool_depth = pool_depth_mean
            ),
          list(
            mean = ~mean(., na.rm = T),
            sd = ~sd(., na.rm = T),
            min = ~min(., na.rm = T),
            max = ~max(., na.rm = T)
          )), .by = "site") 
    , by = "site"
  ) |> 
  left_join(
    hab$lwd_site_year |>
      group_by(site) |> 
      complete(year = min(hab$site_meta$year_min):max(hab$site_meta$year_max)) |> 
      ungroup() |> #glimpse() #no Inf
      summarise(
        lgth_mean = mean(up_down_lgth, na.rm = T),
        lgth_wdth_mean = mean(lgth_wdth, na.rm = T),
        across(
          c(
            lwd_all, lwd_all_per100, lwd_all_per_m2,
            vlwd = lwd_no10, 
            vlwd_per100 = lwd_no10_per100,
            vlwd_per_m2 = lwd_no10_per_m2,
            vlwd_vol
          ),
          list(
            mean = ~mean(., na.rm = T),
            sd = ~sd(., na.rm = T),
            min = ~min(., na.rm = T),
            max = ~max(., na.rm = T)
          )), .by = "site")
    , by = "site"
  ) |>
  #mostly already clean, but coerce a few bankfull_W_D Infs to NA
  mutate(across(where(is.numeric), \(x){ x[is.infinite(x)] <- NA; return(x)})) 

#### 7. measures by site-year for included sites
#keeping only per-year means across transects where applicable
#and again completing site-year cases for explicit missing during full study duration
#and coercing Inf to NA
hab$site_year <- left_join(
  hab$transects_site_year,
  hab$lwd_site_year |> select(-bnk_wdth_mean),
  by = c("strm","site","year","reach_survey_id")
  ) |> 
  semi_join(hab$incl_sites, by = "site") |> 
  select(
    cmplx_strm, strm, site, year, reach_survey_id, n_tsct, n_supp, pct_supp, 
    bankfull_width = bnk_wdth_mean,
    wetted_width = wet_wdth_mean,
    bankfull_W_D = bnk_w_bnk_d_mean,
    wetted_W_D = wet_w_thl_d_mean,
    pct_gravel, 
    pct_pools, 
    pool_depth = pool_depth_mean,
    up_down_lgth, lgth_wdth,
    lwd_all, lwd_all_per100, lwd_all_per_m2,
    vlwd = lwd_no10, 
    vlwd_per100 = lwd_no10_per100,
    vlwd_per_m2 = lwd_no10_per_m2,
    vlwd_vol
  ) |> #summary()
  group_by(cmplx_strm, strm, site) |> 
  complete(year = min(hab$site_meta$year_min):max(hab$site_meta$year_max)) |> 
  ungroup() |> #summary() #glimpse()
  mutate(across(where(is.numeric), \(x){ x[is.infinite(x)] <- NA; return(x)})) 


#### 8. site-year measures made spatial 
#but reduced to only sampled cases due to need for lon/lat
hab$sf_site_year <- hab$site_year |>
  left_join(
    emap$reaches_sampled |> select(site, year, x_lon, x_lat), 
    by = c("site","year")
  ) |> 
  drop_na(x_lon) |> 
  sf::st_as_sf(coords = c("x_lon", "x_lat"), crs = sf::st_crs(4326)) |> 
  sf::st_transform(sf::st_crs(2927)) #NAD83 HARN WA ft


#### final 'clean' objects broken from list for EDA and fitting
#left as extra step to allow last minute wrangling while preserving upstream hab-list versions
site_meta <- hab$site_meta 
sf_site_meta <- site_meta |>
  st_as_sf(coords = c("x_lon_mean", "x_lat_mean"), crs = sf::st_crs(4326)) |>
  sf::st_transform(sf::st_crs(2927))
site_year <- hab$site_year
sf_site_year <- hab$sf_site_year
#whole-stream per-year distributional moments across sites[year]
#distinct from per-site summaries across years in hab$site_meta
#declaration begins from already-site-filtered object
strm_year <- site_year |>  #as_tibble(sf_site_year) |> 
  summarise(
    n_site_with_bnk = length(na.omit(bankfull_width)),
    across(
      c(
        bankfull_width,
        bankfull_W_D,
        pct_gravel, 
        pct_pools, 
        pool_depth,
        up_down_lgth,
        lwd_all, lwd_all_per100, lwd_all_per_m2,
        vlwd, vlwd_per100, vlwd_per_m2,
        vlwd_vol
        ),
      list(
        min = ~min(., na.rm = T),
        q25 = ~quantile(., p = 0.25, na.rm = T),
        q50 = ~median(., na.rm =T),
        mean = ~mean(., na.rm = T),
        q75 = ~quantile(., p = 0.75, na.rm = T),
        max = ~max(., na.rm = T)
      )
    )
    , .by = c("cmplx_strm", "year")
  )

rm(emap)

save.image(file = paste0("~/T/DFW-Team WDFW Watershed Synthesis - IMW analyses/imw_hab_",str_remove_all(Sys.Date(),"-") |> str_sub(3,8),".RData"))

```

Observations were first summarized across transects per site per year. Average annual site pool depth was calculated as the arithmetic mean of the deepest (thalweg) depth measured at transects denoted as intersecting pools. The proportion of gravel was calculated by dividing the number of points classed as coarse or fine gravel by the total count of substrate points (e.g., per site-year, typically 21 transects with 5 points per transect). The density of very large wood was calculated as the counts of pieces >= 0.3m diameter and at least 5m in length, normalized by the total surveyed length between site transects and multiplied by 100. The proportion of pools was the number of transects intersecting a channel pool form, divided by the total number of transects per site per year, and the average width to depth ratio was the mean of per transect bankfull channel width divided by thalweg depth.

Following extensive exploratory analysis, these site-year measures were then assessed for patterns of variation through time, among sites within streams and across streams and complexes. As indicators of 'whole-stream' habitat condition, the per year mean across sites within each stream was also calculated, and these stream-year averages were examined for patterns of change. Mann-Kendall (MK) non-parametric trend tests are a robust and well-established approach to detecting consistent directional change (monotonic increase or decrease) without assumptions of normality in the sample distribution. This simple method was chosen for presentation after results showed qualitative agreement with various other statistical approaches that require additional assumptions and parameter interpretation. The Sen's slope and statistical significance estimate offer insight into questions such as, "Did the proportion of gravel substrate throughout sampled sites in the stream network increase over the study duration?" To account for the temporal autocorrelation evident in many of the time series of yearly whole-stream averages, all trend calculations used the modified MK test in [`rtrend::mkTrend`](https://rpkgs.github.io/rtrend/) (Kong and Song 2024), which implements the approach of Hamed and Rao (1998).


# Results & Discussion

Across complexes, the range through time in habitat measures varied considerably among sites throughout each stream network (Figure SVN). 

Indeed, ranges in the proportion of gravel substrate differed as much or more between sites within streams as between streams (Table SVN). All streams included sites with maximum ranges greater than 0.5, and 7 of 10 streams had much less variable sites with ranges < 0.2, including 0 or nearly 0 difference through time in Stavis, Big Beef, Mill and Abernathy creeks. The per-stream average ranges fell between 0.33 (Stavis) and 0.5 (Germany), with the Lower Columbia streams grouped relatively closely, and the Straits, Little Anderson and Seabeck Creeks also grouped around an average range of approximately 0.4. In other words, all streams included a preponderance of sites with modest change through time in the proportion of gravel substrate as well as several locations characterized by little to no change and several with very substantial variation.

Similarly, site differed as much or more within streams as between streams in terms of the range through time in average pool depth. Ranges of 0 or nearly 0 were present in all streams, and the difference among streams in average range, from 0.2 in Little Anderson Creek to 0.5 in Abernathy Creek, was less than the difference between minimum and maximum ranges in all streams. That is, every stream included locations with pool depths that exhibited very little fluctuation as well as those that showed major change through time. In terms of the absolute magnitude of difference in measured depth, the degree of temporal variation was related as much to position within the stream network as it was to which stream network.

Per-site large wood density showed somewhat greater separation among complexes and streams, although each stream included high and low variation sites. The maximum and average ranges in the Straits streams were larger than in Hood Canal or the Lower Columbia, although every stream other than Mill Creek contained sites with ranges as large or larger than the Straits average. 

Despite the basin-scale differences in geologic and land use history, as well as the contemporary differences between streams in land cover and restoration effort, these long term monitoring data showed streams were consistent in having within-network spatial variation in the magnitude of temporal variation, a finding that clearly reinforces the concept of 'process domains' within riverscapes (e.g., Vannote et al. 1980, Montgomery 1989, Poole 2002). For the practice of process-based river restoration (Beechie et al. 2010), the common occurrence across systems of locations with different levels of change through time underscores the importance of attention to network position and localized dynamics in the diagnosis of limiting factors and the implementation of treatments intended to alleviate them.

```{r figSVN_site_range_ridges, fig.cap="Figure SVN: Stream network variation in temporal variation of habitat measures. Dots are range (max - min) for individual sites; lower interval bar shows IQR and among-site median (diamond). Asterisks designate 'control' streams in original study design."}
d <- site_meta |> #glimpse()
  #filter(year_n > 16) |> 
  select(cmplx_strm, site, 
         ends_with("min"), ends_with("max"), 
         -starts_with("year"), 
         -contains("lwd_all"), #only 'very' large wood
         -vlwd_min, -vlwd_max #use density not counts
         , -contains("bankfull_width")
         , -contains("wetted")
         , -contains("_m2")
         ) |> 
  pivot_longer(-c(cmplx_strm, site)) |> 
  mutate(
    m = str_sub(name, -3,-1), # m = map_chr(meas, ~last(str_split_1(.x,"_"))),
    name = str_remove(name, paste0("_",m)) # meas = map2_chr(meas, m, ~str_remove(.x, paste0("_",.y)))
  ) |> 
  pivot_wider(names_from = m, values_from = value) |> 
  mutate(
    site_rng = max - min,
  ) |> 
  drop_na(site_rng) |> 
  filter(is.finite(site_rng))

d |> 
  ggplot(mapping = aes(site_rng, fct_rev(cmplx_strm), color = cmplx_strm, fill = cmplx_strm)) +
  geom_vline(xintercept = 0) +
  ggridges::geom_density_ridges(
    scale = 0.9,
    alpha = 0.6,
    show.legend = F) +
  ggdist::stat_dots(
    #this would work in a split/map pattern but with faceting here: aes(slab_alpha = after_stat(x)),
    orientation = "horizontal",
    alpha = 0.6,
    show.legend = F
  ) +
  ggdist::stat_pointinterval(
    orientation = "horizontal",
    color = "black",
    .width = c(0.5, 0.95),
    shape = 3, #18,
    fatten_point = 3,
    show.legend = F
  ) +
  coord_cartesian(xlim = c(0, NA), ylim = c(0.7, NA), expand = F) +
  scale_color_manual(values = pal_cmplx_strm, aesthetics = c("color","fill")) +
  facet_wrap(~name, 
             nrow = 1, #ncol = 3,
             scales = "free_x") +
  labs(
    x = "", y = ""
    ,
    title = "Stream network variation in range of habitat measures through time",
    subtitle = "Dots are range (max - min) for individual sites, interval shows median, IQR and 95% span"
  )

ggsave(filename = "fig_SVN_site_range_ridges.png", width = 11, height = 9, dpi = 150, bg = "white")
```

```{r tabSNV_site_range_ridges, tab.cap="Table SNV: Mean, minimum, and maximum range per-site in habitat measures across streams."}
d_gt <- d |> 
  summarise(
    across(
      #c(min, max, site_rng), 
      site_rng,
      list(
        min = ~min(.,na.rm = T),
        mean = ~mean(.,na.rm = T),
        max = ~max(.,na.rm = T)
        )
      )
    ,.by = c(cmplx_strm, name)
  ) |>
  left_join(
    distinct(site_meta, cmplx_strm, cmplx_desc), by = "cmplx_strm"
  ) |> 
  pivot_wider(names_from = name, values_from = contains("site_rng")) |> 
  select(
    cmplx_desc, cmplx_strm, #strm,
    contains("width"), contains("W_D"), contains("gravel"), contains("pools"), contains("depth"), contains("vlwd")
  )

d_gt |> 
  gt(
    groupname_col = "cmplx_desc", 
    rowname_col = "cmplx_strm", 
    caption = "Mean, minimum, and maximum range per-site in habitat measures across streams"
    ) |> 
  fmt_number(where(is.numeric), decimals = 0) |> 
  fmt_number(contains("pool_depth"), decimals = 1) |> 
  fmt_number(contains("pct"), decimals = 2) |> 
#  tab_spanner("Bankfull W", contains("width")) |> 
  tab_spanner("BnkW:D", contains("W_D")) |> 
  tab_spanner("Ppn. gravel", contains("pct_gravel")) |> 
  tab_spanner("Ppn. pools", contains("pct_pool")) |> 
  tab_spanner("Pool depth", contains("pool_depth")) |> 
  tab_spanner("VLWD/100m", contains("per100")) |> 
  tab_spanner("VLWD vol.", contains("vol")) |> 
  cols_label_with(contains("min"), ~"min") |> 
  cols_label_with(contains("mean"), ~"mean") |> 
  cols_label_with(contains("max"), ~"max") |> 
  tab_style(
    style = list(cell_borders("left", weight = px(1.5))), 
    locations = cells_body(columns = contains("min"))
  ) |> 
  tab_style(
    style = list(cell_text(style = "italic")), 
    locations = map(3:17, 
      ~cells_body(columns = .x, rows = which.min(unlist(d_gt[,.x])))
    )
  ) |> 
  tab_style(
    style = list(cell_text(decorate = "underline")), 
    locations = map(3:17, 
      ~cells_body(columns = .x, rows = which.max(unlist(d_gt[,.x])))
    )
  )

```

In addition to the shared pattern of within-network heterogeneity in the magnitude of temporal variation, consistent directional change at the watershed scale was rarely detectable in the time series of habitat measures summarized across sites per year to a 'whole-stream' average (Table MK-strm-year, Figure MK-stream-year). Furthermore, with the exception of large wood in Abernathy Creek, the detectable trends were weak, failed to indicate separation between treatment and control streams, and did not necessarily suggest in-channel habitat improvement.

Mann-Kendall (MK) tests for trends in whole-stream interannual change were significant (p < 0.05) for the proportion of gravel substrate in Deep and East Twin Creeks (weakly positive slopes indicating increased gravel prevalence), for mean pool depth in Little Anderson and Deep Creeks (weakly negative slopes indicating shallower pools on average), and for large wood density in Abernathy Creek (modestly strong positive slope indicating increased wood prevalence).

Various individual sites did show statistically significant directional shifts through time in several measures (Figure MK-site-year). However, again with the exception of Abernathy Creek's large wood density change, these findings reinforced the conclusions at the scale of the whole-stream averages: relatively weak trends, little treatment-control separation, and little indication of desired habitat improvement.

*[application to practice: restoration effects were only detectable when implemented at sufficient magnitude and duration and extent (i.e., lots of wood, over several years, in enough stream km above monitoring sites)]*


```{r tabMK_strm_year, tab.cap="Mann-Kendall trend test results for interannual change in habitat measures, corrected for temporal autocorrelation. "}
#https://vsp.pnnl.gov/help/vsample/design_trend_mann_kendall.htm

strm_year_mk <- strm_year |> #glimpse()
  select(
    cmplx_strm, year, ends_with("_mean"),
    -contains("lgth"), -contains("_m2"), -contains("bankfull_width"),
    -lwd_all_mean, -lwd_all_per100_mean, -vlwd_mean
    ) |> 
  # rename_with(~str_replace(., "_mean", "_stream_year_mean")) |> 
  rename_with(~str_remove(., "_mean")) |> 
  summarise(
    across(
        where(is.double),
        list(
          mk = ~list(rtrend::mkTrend(.))
          )
        ),
      .by = c(cmplx_strm)
    ) |> 
  unnest_wider(ends_with("_mk"), names_sep = ":") |> 
  select(cmplx_strm, ends_with("pval"), ends_with("slp"))

strm_year_mk |> 
  arrange(cmplx_strm) |> 
  left_join(distinct(site_meta, cmplx_strm, cmplx_desc), by = "cmplx_strm") |> 
  select(cmplx_desc, cmplx_strm, 
         ends_with("slp"), ends_with("pval"), #so that estimate is left of pval
         contains("width"), contains("W_D"), contains("gravel"), contains("pools"), contains("depth"), contains("per100"), contains("vol") #so that measures are side by side
         ) |> 
  gt(
    groupname_col = "cmplx_desc", 
    rowname_col = "cmplx_strm"
    #,caption = "Mann-Kendall trend test results for interannual change in habitat measures"
    ) |> 
  fmt_number(ends_with("pval"), decimals = 2) |> 
  fmt_number(ends_with("slp"), decimals = 4) |> 
#  tab_spanner("Bankfull W", contains("width")) |> 
  tab_spanner("BnkW:D", contains("W_D")) |> 
  tab_spanner("Ppn. gravel", starts_with("pct_gravel")) |> 
  tab_spanner("Ppn. pools", starts_with("pct_pool")) |> 
  tab_spanner("Pool depth", starts_with("pool_depth")) |> 
  tab_spanner("VLWD/100m", contains("per100")) |> 
  tab_spanner("VLWD vol.", contains("vol")) |> 
  cols_label_with(ends_with("slp"), ~"Sen's") |> 
  cols_label_with(ends_with("pval"), ~"p-val") |> 
  tab_style(style = cell_borders("left"), locations = cells_body(columns = ends_with("slp"))) |> 
  tab_style_body(
    style = list(cell_text(color = "red")), 
    columns = c(ends_with("slp"), ends_with("pval")),
    fn = \(x) x < 0 
  ) |> 
  tab_style_body(
    style = list(cell_fill("grey80"), cell_text(weight = "bold")), 
    columns = c(ends_with("pval")),
    fn = \(x) x <= 0.05 
  )
```

```{r figMK_strm_year, fig.cap="Time series of whole-stream average measures with Mann-Kendall test p-value < 0.05. "}
strm_year_mk |> 
  pivot_longer(-cmplx_strm) |> 
  separate(name, into = c("name","stat"), sep = "_mk:") |> 
  pivot_wider(names_from = stat, values_from = value) |> 
  filter(pval < 0.05) |>
  mutate(cmplx_strm = as.character(cmplx_strm)) |> 
  arrange(name) |> 
  mutate(
    gg = map2(
      name, cmplx_strm,
      ~strm_year |> 
        filter(cmplx_strm == .y) |> 
        select(cmplx_strm, year, yval = all_of(paste0(.x,"_mean"))) |> 
        ggplot(aes(year, yval, color = cmplx_strm)) + 
        geom_line(linetype = 2, show.legend = F) + 
        geom_point(show.legend = F) + 
        #geom_smooth(se = F, show.legend = F) + 
        scale_color_manual(values = pal_cmplx_strm) +
        labs(x = "", y = .x, subtitle = .y)
    )
  ) |> 
  pull(gg) |> 
  wrap_plots(ncol = 4) #+plot_annotation(title = "Time series of measures with Mann-Kendall test p-value < 0.05")

#ggsave(filename = "fig_MK_strm_year.png", width = 11, height = 9, dpi = 150, bg = "white")
```

\n

```{r figMK_site_year, fig.cap="Individual site very large wood density through time. Colored series indicate sites with significant (p < 0.05) Mann-Kendall trends, overlaid on the remaining sites (light gray)."}
site_year_mk <- site_year |> #glimpse()
  select(
    cmplx_strm, site, year, 
    bankfull_W_D, pct_gravel, pct_pools, pool_depth, vlwd_per100
    ) |> 
  summarise(
    across(
        where(is.double),
        list(
          mk = ~list(rtrend::mkTrend(.))
          )
        ),
      .by = c(cmplx_strm, site)
    ) |> 
  unnest_wider(ends_with("_mk"), names_sep = ":") |> 
  select(cmplx_strm, site, ends_with("pval"), ends_with("slp")) |> 
  pivot_longer(-c(cmplx_strm, site))

site_year |> 
  select(
    cmplx_strm, site, year, 
    bankfull_W_D, pct_gravel, pct_pools, pool_depth, vlwd_per100, vlwd_vol
  ) |> 
  pivot_longer(-c(cmplx_strm, site, year)) |> 
  #semi_join(
  left_join(
    site_year_mk |> 
      select(-cmplx_strm, site, name, pval = value) |> 
      filter(str_detect(name, "pval")) |>  #1345 site-measure tests
      filter(pval < 0.05) |> #166 'signif'
      #count(name) |> arrange(desc(n)) #most for wood
      #count(cmplx_strm) |> arrange(desc(n)) #most for Abernathy; note Deep and Stavis very similar to treatment streams in n-signif
      mutate(
        name = str_remove(name, "_mk:pval")
        )
    ,
    by = c("site","name")    
  ) |> 
  filter(name == "vlwd_per100") |> 
  split(~name) |> 
  map(
    ~ggplot(mapping = aes(year, value, group = site, color = cmplx_strm)) +
      geom_line(data = .x |> filter(is.na(pval)), color = "grey", linewidth = 0.2, show.legend = F) +
      geom_line(data = .x |> drop_na(pval), linewidth = 1.1, show.legend = F) +
      geom_point(data = .x |> drop_na(pval), size = 0.3, show.legend = F) +
      scale_color_manual(values = pal_cmplx_strm) +
      facet_wrap(~cmplx_strm, scales = "free") +
      labs(x = "", y = "", subtitle = .x$name[1])
  )

ggsave(filename = "fig_MK_site_year.png", width = 11, height = 9, dpi = 150, bg = "white")
```


**Possible further points (delete/modify as desired)**

  - unanticipated limitations of the original treatment-control study design: not enough concentrated 'treatment' interventions because of limited alignment with lead entity priorities and inadequately weighted competition with non-IMW projects

  - We do not see differentiation relative to 'treatment-control' across streams within complex, but there is evidence of non-linear longer-phase cycles that could be viewed as multi-year 'trends' (GAM regressions on smoothed year). This has bearing on restoration practice, both for planning (e.g., shifts in perceived limiting factors due to large scale hydroclimate fluctuation) and effectiveness monitoring (e.g., risk of both/either false assertion/attribution of trend and/or incorrect assumption that work is not working).

# Conclusions

Successful salmon recovery requires thinking upstream and down, next year and the year after and five years after that.

# Cited

Anlauf, K.J., Jones, K.K., and Stein, C.H. 2009. The status and trend of physical habitat and rearing potential in coho bearing streams in the Oregon coastal coho evolutionary significant unit. Oregon Department of Fish and Wildlife, Salem, Ore.

Beechie, T.J., Sear, D.A., Olden, J.D., Pess, G.R., Buffington, J.M., Moir, H., Roni, P. and Pollock, M.M., 2010. Process-based principles for restoring river ecosystems. BioScience, 60(3), pp.209-222.

Bilby, R.E., Currens, K.P., Fresh, K.L., Booth, D.B., Fuerstenberg, R.R. and Lucchetti, G.L., 2024. Why Aren't Salmon Responding to Habitat Restoration in the Pacific Northwest?. Fisheries, 49(1), pp.16-27.

Crawford, B.A., Drivdahl, C., Leider, S., Richmond, C. and Butkus, S., 2002. The Washington Comprehensive Monitoring Strategy for Watershed Health and Salmon Recovery.

Hamed, K. H., & Rao, A. R., 1998. A modified Mann-Kendall trend test for autocorrelated data. Journal of hydrology, 204(1-4), 182-196. <doi:10.1016/S0022-1694(97)00125-X>.

Kong D, Song H, 2024. _rtrend: Trend Estimating Tools_. R package version 0.1.5, <https://CRAN.R-project.org/package=rtrend>.

Lazorchak, J., D. Klemm, and D. Peck, Editors. (1998). Environmental Monitoring and Assessment Program-Surface Waters: Field Operations and Methods for Measuring the Ecological Condition of Wadeable Streams.

Montgomery, D.R., 1999. Process domains and the river continuum 1. JAWRA Journal of the American Water Resources Association, 35(2), pp.397-410.

Parker, K.R. and Wiens, J.A., 2005. Assessing recovery following environmental accidents: environmental variation, ecological assumptions, and strategies. Ecological Applications, 15(6), pp.2037-2051.

Poole, G.C., 2002. Fluvial landscape ecology: addressing uniqueness within the river discontinuum. Freshwater biology, 47(4), pp.641-660.

Roni, P., editor. 2005. Monitoring stream and watershed restoration.  American Fisheries Society, Bethesda, Maryland.

Roper, B.B., J.L. Kershner, and R.C. Henderson. 2003. The value of using permanent sites when evaluating stream attributes at the reach scale.  Journal of Freshwater Ecology  18(4): 585 – 592.

Stevens, D.L., Jr., and A.R. Olsen.  2004.  Spatially balanced sampling of natural resources.  Journal of the American Statistical Association 99: 262-278.

Vannote, R. L., Minshall, G. W., Cummins, K. W., Sedell, J. R., and Cushing, C. E., 1980. The river continuum concept. Canadian Journal of Fisheries and Aquatic Sciences, 37(1), 130-137

# Supplementary

```{r fig_strm_year_all, eval=FALSE}
strm_year |> 
  select(
    cmplx_strm, year, ends_with("_mean"),
    -contains("lgth"), -contains("_m2"), -contains("bankfull_width"),
    -lwd_all_mean, -lwd_all_per100_mean, -vlwd_mean
  ) |> 
  # rename_with(~str_replace(., "_mean", "_stream_year_mean")) |> 
  rename_with(~str_remove(., "_mean")) |> 
  pivot_longer(cols = -c(cmplx_strm, year)) |> 
  ggplot(aes(year, value, color = cmplx_strm)) + 
  geom_line(linetype = 2, show.legend = F) + 
  geom_point(size = 0.7, show.legend = F) + 
  #geom_smooth(se = F, show.legend = F) + 
  scale_color_manual(values = pal_cmplx_strm) +
  facet_wrap(~name + cmplx_strm, ncol = 10, scales = "free") +
  labs(x = "", y = "", subtitle = "Time series of whole stream mean across sites per year")

ggsave(filename = "fig_strm_year_all.png", width = 16, height = 12, dpi = 150, bg = "white")
```


```{r gg_site_range_horiz_cascade, eval=FALSE}
d |>
  split(~name) |>
  map(
    ~.x |> 
      drop_na(min) |> 
      ggplot() +
      geom_linerange(
        aes(xmin = min, xmax = max,
            y = fct_reorder(site, min, .fun = "min"),
            color = cmplx_strm)
        , show.legend = F
      ) +
      scale_color_manual(values = pal_cmplx_strm, aesthetics = c("color","fill")) +
      labs(y = "", subtitle = .x$name[1])
  )  # wrap_plots(nrow = 1)

```

```{r gg_strm_year_pointrange_series_grid, eval=FALSE}
strm_year |> #glimpse()
  select(cmplx_strm, year,
         ends_with("min"), ends_with("mean"), ends_with("max"), 
         -contains("lwd_all"), 
         -vlwd_min, -vlwd_mean, -vlwd_max #use density not counts 
         ,-contains("up_down"),-contains("bankfull_width")
  ) |> 
  pivot_longer(-c(cmplx_strm, year)) |> 
  mutate(
    m = str_sub(name, -3,-1),
    m = if_else(m == "ean","mean",m), #quickfix 
    name = str_remove(name, paste0("_",m)) 
  ) |> 
  pivot_wider(names_from = m, values_from = value) |> 
  ggplot(aes(year, mean, color = cmplx_strm)) +
  geom_line(linetype = 2, linewidth = 0.5, show.legend = F) +
  geom_linerange(aes(ymin = min, ymax = max), linewidth = 0.5, show.legend = F) +
  #geom_linerange(aes(ymin = q25, ymax = q75), linewidth = 1.2, show.legend = F) +
  geom_point(show.legend = F) +
  scale_color_manual(values = pal_cmplx_strm) +
  facet_grid(cols = vars(cmplx_strm), rows = vars(name), scales = "free") +
  labs(
    x = "", y = "",
    title = "Whole-stream per-year means across sites"
  )

ggsave(filename = "fig_strm_year_pointrange_series_grid.png", width = 16, height = 12, dpi = 150, bg = "white")
```

```{r strm_year_ts_kpss, eval=FALSE}
# library(tsibble)
# library(feasts) #loads: library(fabletools)
# library(fable)
#https://feasts.tidyverts.org/

strm_year_ts <- d |> 
  tsibble::as_tsibble(key = cmplx_strm, index = year)

# strm_year_ts |> fabletools::features(c(pct_gravel_mean), features = quantile)
# strm_year_ts |> fabletools::features(c(pct_gravel_mean), features = list(quantile, mu = mean))
# strm_year_ts |> fabletools::features(c(pct_gravel_mean), features = feat_acf)
# # # STL (via loess) not appropriate for yearly data, with no meaningful 'season'?
# # #also "trend_strength" is relative to 'seasonal' component, not really "slope magnitude"
# # strm_year_ts |> fabletools::features(c(pct_gravel_mean), features = feasts::feat_stl)
# 
# #gravel MK signif (pos) for Deep and E-twin
# strm_year_ts |> fabletools::features(c(pct_gravel_mean), features = unitroot_kpss) |> arrange(kpss_pvalue)

strm_year_ts_kpss <- strm_year_ts |>
  pivot_longer(-c(cmplx_strm, year)) |> 
  split(~name) |> 
  map(
    ~fabletools::features(.x, value, features = feasts::unitroot_kpss) |> 
      mutate(name = .x$name[1])
      #rename_with(~paste(.x$name[1],.,sep = "_"), starts_with("kpss"))
    ) |> 
  bind_rows()

strm_year_ts_kpss |> 
  # filter(kpss_pvalue < 0.05)
  pivot_wider(names_from = name, values_from = starts_with("kpss")) |> 
  left_join(distinct(site_meta, cmplx_strm, cmplx_desc), by = "cmplx_strm") |> 
  select(cmplx_desc, cmplx_strm, 
         #contains("stat"), contains("pval"), #so that estimate is left of pval
         contains("width"), contains("W_D"), contains("gravel"), contains("pools"), contains("depth"), contains("vlwd") #so that measures are side by side
  ) |> 
  gt(
    groupname_col = "cmplx_desc", 
    rowname_col = "cmplx_strm", 
    caption = "KPSS test results for interannual stationarity in habitat measures"
  ) |> 
  fmt_number(contains("pvalue"), decimals = 2) |> 
  fmt_number(contains("stat"), decimals = 2) |> 
  tab_spanner("Bankfull W", contains("width")) |> 
  tab_spanner("BnkW:ThalD", contains("W_D")) |> 
  tab_spanner("Ppn. gravel", contains("pct_gravel")) |> 
  tab_spanner("Ppn. pools", contains("pct_pool")) |> 
  tab_spanner("Pool depth", contains("pool_depth")) |> 
  tab_spanner("VLWD/100m", contains("vlwd")) |> 
  cols_label_with(contains("stat"), ~"Stat.") |> 
  cols_label_with(contains("pval"), ~"p-val") |> 
  tab_style(style = cell_borders("left"), locations = cells_body(columns = contains("stat"))) |> 
  tab_style_body(
    style = list(cell_text(color = "red")), 
    columns = c(contains("stat"), contains("pvalue")),
    fn = \(x) x < 0 
  ) |> 
  tab_style_body(
    style = list(cell_fill("grey80"), cell_text(weight = "bold")), 
    columns = c(contains("pvalue")),
    fn = \(x) x <= 0.05 
  )

```

```{r dew765, eval=FALSE}
site_year |> filter(site == "DEW765") |> ggplot(aes(year, vlwd_per100)) + geom_line() + geom_point()

sf_site_year |> 
  filter(site == "DEW765") |> 
  ggplot() +
  #geom_sf_text(data = sf_site_meta |> filter(site == "DEW765"), aes(label = site)) +
  geom_sf(data = sf_site_year |> filter(site == "DEW765") |> select(site), color = "grey", alpha = 0.5) +
  geom_sf(aes(size = vlwd)) +
  facet_wrap(~year)

```

